Parallel air palletization and tour planning for simultaneous pickup and delivery

### 2022-11-10 latex/comp_or.tex
Paper updated

\title{Multiprocessing air palletization and tour planning for simultaneous pickup and delivery}

@@

In the scenario of aerial pickup and delivery of goods in a distribution network, transport aviation faces risks of cargo unbalancing due to the urgency required for loading for rapid take-off and mission accomplishment, especially in times of crisis, public calamity, client contract short lead times, or any external pressure for immediate take-off. Also, there is no technological assistance to help the load and trip planners with the huge number of demands for transport in each hub.

This enables other risks such as improper delivery, excessive fuel burn, and longer than necessary turn-around time. We contributed by modeling and solving a new problem of planning the loading and routing of an aircraft according to a utility score, weight and balance principles, and fuel consumption in a tour of simultaneous pickup and delivery at intermediate hubs.

This new hard problem, named {\it Air Cargo Load Planning with Routing, Pickup, and Delivery Problem (ACLP+RPDP)}, is mathematically modeled using standardized pallets in fixed positions, obeying the center of gravity constraints, delivering each item to its destination, and minimizing fuel consumption costs.

We also contributed by carrying out multiple experiments with a commercial solver and the well-known Ant Colony Optimization on synthetic data based on real data from the {\it Brazilian Air Force}\/ transportation history.

These challenging benchmark instances are made publicly available. We also created a multiprocessing computing heuristic that quickly finds good solutions for a wide range of problem sizes, an essential contribution as it was the unique method that managed to solve all testing scenarios.

@@

On the other hand, our tactic will be, at each shipping node, to predefine the destinations of the pallets at that node. In this way, we will reserve a number of pallets proportional to the volume demanded by each destination at the shipping node. We could have used another criterion, but it was observed in the experiments that volume is more constrictive in airlift. Once the destinations of the pallets are defined, we will use heuristics and a MIP solver to find the best possible node-by-node solutions. This strategy is summarized in Algorithm \ref{alg:main}.

@@

$method$\/ corresponds to a MIP solver or one of the heuristics that we will present in subsection \ref{methods}. The loop of lines \ref{main:loop1}-\ref{main:loop2} goes through all permutations $\pi$, where the node-by-node resolutions are performed by $SolveTour$, whose result is stored in $f_{\pi}$. The best result among all $K!$\/ tours will be the answer for $scenario$, $volume$\/ and $method$\/ (line \ref{main:f}).

@@

In addition to the set of nodes, pallets, costs and items, $SolveTour$, described in Algorithm \ref{alg:tour}, receives the parameter $method$, which corresponds to a MIP solver or a heuristic for solving the node-by-node problems, and the parameter $\pi$, which is a permutation that defines the order of visits in this tour.

@@

In this subsection we present implementations of $SolveNode(method,k, G)$, where $method$\/ corresponds to a MIP solver or a heuristic, $k$\/ is the index of the current node and $G$\/ is the allocation graph of the consolidated items that remain on board at node $l_k$.

@@

		{\it Heuristic's procedures:}  & All heuristics use the same procedures for selecting and inserting edges.  \\

@@

		{\it Common procedures:}  & All methods have the same procedures for: solution printing, pallets destinations and initial parameters setting, and solution integrity checking.\\

@@

We use four types of random selections:
\begin{itemize}
	\item $RandomReal(r_1,r_2)$: randomly selects a real number in $[r_1,r_2]$;
	\item $RandomInt(i_1,i_2)$: randomly selects a integer number in $[i_1,i_2]$;
	\item $RandomChoice(set)$: randomly selects an element from $set$;
	\item $Roulette(set)$ biased through $\phi$: selects an element from $set$, where the probability of each element is proportional to the value of a given function $\phi$\/ defined on $set$.
\end{itemize}

@@

Below, we present five implementations for $SolveNode$: the first using a MIP solver, and four with developed heuristics: {\it Ant Colony Optimization}\/ (ACO), {\it Multiprocessed Ant Colony Optimization}\/ ({\it mpACO}),  and {\it Shims},  and {\it Multiprocessed Shims (mpShims)}.

@@

\subsubsection{MIP solver}
\label{solver}

To find a viable solution for ACLP+RPDP, the strategy adopted in $SolveTour$\/ previously defines the values of some variables. Concretely, all permutations between the nodes are tested, the set of nodes to be visited is updated, the consolidated items that remain on board are reallocated to minimize the CG deviation, and the pallets destination is determined according to the volume of items available for shipment. In this way, the mathematical modeling for $SolveNode(MIP,k,G)$\/ becomes simpler, which finds an allocation of available items in node $l_k$\/ using previously defined values of $L_k$, $p_i.to[k]$, $Y^k_{iq}$\/ and $a^k_q$.

This simplified modelling is described in the equations \ref{eq:newf} to \ref{eq:lastnew}, which maximizes the score of the $t^k_j$\/ items to be shipped, maintaining load balancing. The variables $X^k_{ij}$\/ define the set of edges $E^N_k$, which will be added to the graph $G$.

\begin{equation} \label{eq:newf}
	\max f_s = \sum_{i=1}^{m} \sum_{j=1}^{n_k} X_{ij}^k \times t_j^k.s
\end{equation}

\begin{equation} 
	maxW = \min(Payload, \sum_{i=1}^{m}p_i.w)
\end{equation}

\begin{equation} 
	\epsilon_k = \sum_{i=1}^{m}[ p_i.d \times (\sum_{j=1}^{n_k} X_{ij}^k \times t_j^k.w +  \sum_{q=1}^{m} Y_{iq}^k \times a_q^k.w)]/(maxW \times limit^{CG}_{long})
\end{equation}

\begin{equation}
	X_{ij}^k = 0 \ \mbox{if} \ t_j^k.to \notin L_k; \ i \in \{1, 2, \ldots, m\}; \ j \in \{1, 2, \ldots, n_k\}
\end{equation}

\begin{equation} 
	LatIt_k = \sum_{i=1}^{m} \sum_{j=1}^{n_k} ( X_{ij}^k \times t_j^k.w \times (i\%2) - X_{ij}^k \times t_j^k.w \times (i+1)\%2 )
\end{equation}

\begin{equation} 
	LatCons_k =  \sum_{i=1}^{m} \sum_{q=1}^{m_k}  ( Y_{iq}^k \times a_q^k.w \times (i\%2) - Y_{iq}^k \times a_q^k.w \times (i+1)\%2)
\end{equation}

\begin{equation} 
	s.t.: d^{CG}_{pallet} \times | LatIt_k + LatCons_k | \leq  \sum_{i=1}^{m}p_i.w \times limit^{CG}_{lat}
\end{equation}

\begin{equation}
	s.t.: |\tau_k| \leq maxW \times limit^{CG}_{long}
\end{equation}

\begin{equation} 
	s.t.: \sum_{i=1}^{m} (\sum_{j=1}^{n_k} X_{ij}^k \times t_j^k.w + \sum_{q=1}^{m_k} Y_{iq}^k \times a_q^k.w ) \leq maxW
\end{equation}

\begin{equation} 
	s.t.: \sum_{j=1}^{n_k} X_{ij}^k \times t_j^k.w + \sum_{q=1}^{m_k} Y_{iq}^k \times a_q^k.w  \leq p_i.w; \ i \in \{1, 2, \ldots, m_k\}
\end{equation}

\begin{equation} 
	s.t.: \sum_{j=1}^{n_k} X_{ij}^k \times t_j^k.v + \sum_{q=1}^{m_k} Y_{iq}^k \times a_q^k.v  \leq\ p_i.v; \ i \in \{1, 2, \ldots, m_k\}
\end{equation}

\begin{equation} 
	s.t.: \sum_{i=1}^{m} X_{ij}^k \leq 1; \ j \in \{1, 2, \ldots, n_k\}
\end{equation}

\begin{equation} \label{eq:lastnew}
	s.t.: p_i.to[k] = t^k_j.to\ \mbox{if} \ X_{ij}^k = 1; \ i \in \{1, 2, \ldots, m\}; \ j \in \{1, 2, \ldots, n_k\}
\end{equation}

@@

For parallelization, we used the {\it Python multiprocessing} package, which implements {\it process-based parallelism}. According to https://docs.python.org/3/library/multiprocessing.html, Multiprocessing is a package that supports spawning processes using an API similar to the threading module. The multiprocessing package offers both local and remote concurrency, effectively side-stepping the global interpreter lock by using sub-processes instead of threads. Because of this, the multiprocessing module lets the programmer take full advantage of the fact that a machine has more than one processor. This decision for multiprocessing relies on the need to exchange information among processes, which would not be possible with thread-based parallelism.

Process: The operating system’s spawned and controlled entity that encapsulates an executing application. A process has two main functions. The first is to act as the resource holder for the application, and the second is to execute the instructions of the application.

— PAGE 271, THE ART OF CONCURRENCY, 2009. The Art of Concurrency: A Thread Monkey's Guide to Writing Parallel Applications 1st Edition, Kindle Edition by Clay Breshears  (Author)

parallel shared memory algorithms

>>>>

\title{Parallel air palletization and tour planning for simultaneous pickup and delivery}

@@

In the scenario of aerial pickup and delivery of goods in a distribution network, transport aviation faces risks of cargo unbalancing due to the urgency required for loading for rapid take-off and mission accomplishment, especially in times of crisis, public calamity, client contract short lead times, or any external pressure for immediate take-off. Also, there is no technological assistance to help the load and trip planners with the huge number of demands for transport in each hub.
This enables other risks such as improper delivery, excessive fuel burn, and longer than necessary turn-around time. We contributed by modeling and solving a new problem of planning the loading and routing of an aircraft according to a utility score, weight and balance principles, and fuel consumption in a tour of simultaneous pickup and delivery at intermediate hubs.
This new hard problem, named {\it Air Cargo Load Planning with Routing, Pickup, and Delivery Problem (ACLP+RPDP)}, is mathematically modeled using standardized pallets in fixed positions, obeying the center of gravity constraints, delivering each item to its destination, and minimizing fuel consumption costs.
We also contributed by carrying out multiple experiments with the well-known Ant Colony Optimization on synthetic data based on real data from the {\it Brazilian Air Force}\/ transportation history.
These challenging benchmark instances are made publicly available. We also created a multiprocessing computing heuristic that quickly finds good solutions for a wide range of problem sizes, an essential contribution as it was the unique method that managed to solve all testing scenarios.

@@

On the other hand, our tactic will be, at each shipping node, to predefine the destinations of the pallets at that node. In this way, we will reserve a number of pallets proportional to the volume demanded by each destination at the shipping node. We could have used another criterion, but it was observed in the experiments that volume is more constrictive in airlift. Once the destinations of the pallets are defined, we will use serial and multiprocessing heuristics to find the best possible node-by-node solutions. This strategy is summarized in Algorithm \ref{alg:main}.

@@

$method$\/ corresponds to one of the heuristics that we will present in subsection \ref{methods}. The loop of lines \ref{main:loop1}-\ref{main:loop2} goes through all permutations $\pi$, where the node-by-node resolutions are performed by $SolveTour$, whose result is stored in $f_{\pi}$. The best result among all $K!$\/ tours will be the answer for $scenario$, $volume$\/ and $method$\/ (line \ref{main:f}).

@@

In addition to the set of nodes, pallets, costs and items, $SolveTour$, described in Algorithm \ref{alg:tour}, receives the parameter $method$, which corresponds to a heuristic for solving the node-by-node problems, and the parameter $\pi$, which is a permutation that defines the order of visits in this tour.

@@

In this subsection we present implementations of $SolveNode(method,k, G)$, where $method$\/ corresponds to a heuristic, $k$\/ is the index of the current node and $G$\/ is the allocation graph of the consolidated items that remain on board at node $l_k$.

@@

@@

		{\it Common procedures:}  & All methods have the same procedures for: solution printing, pallets destinations and initial parameters setting, solution integrity checking, and the same procedures for selecting and inserting edges.\\

@@

We use four types of random selections:
\begin{itemize}
	\item $RandomReal(r_1,r_2)$: randomly selects a real number in $[r_1,r_2]$;
	\item $RandomInt(i_1,i_2)$: randomly selects a integer number in $[i_1,i_2]$;
	\item $Roulette(set)$ biased through $\phi$: selects an element from $set$, where the probability of each element is proportional to the value of a given function $\phi$\/ defined on $set$.
\end{itemize}

@@

Below, we present four implementations for $SolveNode$: {\it Ant Colony Optimization}\/ (ACO), {\it Multiprocessed Ant Colony Optimization}\/ ({\it mpACO}),  and {\it Shims},  and {\it Multiprocessed Shims (mpShims)}.

@@

@@

For parallelization, we used the {\it Python multiprocessing} package, which implements {\it process-based parallelism or parallel shared memory algorithm}. According to \cite{multiprocessing}, Multiprocessing is a package that supports spawning processes using an API similar to the threading module.

The multiprocessing package offers both local and remote concurrency, effectively side-stepping the global interpreter lock by using sub-processes instead of threads. Because of this, the multiprocessing module lets the programmer take full advantage of the fact that a machine has more than one processor. This decision for multiprocessing relies on the need to exchange information among processes, which would be harder to handle with thread-based parallelism.

According to \cite[p.271]{Breshears2009}, a Process is the operating system’s spawned and controlled entity that encapsulates an executing application. A process has two main functions: The first is to act as the resource holder for the application, and the second is to execute the instructions of the application.
